Robust Control of a Fluxonium Qubit

### 2020-11-20 paper/1introduction.tex
minor edits

\section{Introduction}
Quantum optimal control (QOC) techniques are a class of optimization
algorithms for accurately and efficiently manipulating quantum systems.
Early techniques were proposed for nuclear magnetic resonance experiments
\cite{khaneja2005optimal}, and applications now include superconducting
circuits \cite{heeres2017implementing,
  huang2020engineering, leng2019robust, leung2017speedup, xu2020nonadiabatic},
neutral and ionized atoms \cite{van2016optimal}, nitrogen-vacancy centers in
diamond \cite{rembold2020introduction}, and Bose-Einstien condensates
\cite{sorensen2018quantum}. For quantum computation
QOC techniques are employed to achieve high fidelity gates
while adhering to experimental constraints.
The decision variables of the optimization problem are the time-dependent control
parameters, particular to the quantum system, that govern its evolution.
Experimental errors may cause the system evolution to deviate from that predicted in
optimization, hampering performance.
The field of classical control theory has developed robust control techniques
to encode experimental errors in optimization objectives, improving
experimental performance (citations appreciated).
In this work we employ robust control techniques to mitigate
realistic system parameter deviations and decoherence that arise when controlling
a superconducting fluxonium qubit.

Analytic and numerical techniques for QOC have been remarkably successful in
designing high fidelity gates (citations needed). 
While these methods are able to predict system behavior with high accuracy, they tend 
to be sensitive to experimental errors such as parameter drift, noise, and finite control 
resolution. 
% In addition to the high
% simulated fidelities which can be obtained with these techniques, it is desirable for the
% resultant quantum state trajectory to be insensitive to experimental errors
% such as parameter drift, noise, and finite control resolution. Quantum state
% trajectories that are insensitive to these effects are said to be robust.
% In line with this objective, 
In order to address these issues,
recent work has sought to mitigate decoherence
and errors due to parameter deviations
Considering the dynamical and geometric phases of quantum state
trajectories has led to analytic methods for mitigating
errors due to parameter deviations and pure dephasing
\cite{han2020experimental, merrill2014progress, xu2020nonadiabatic, zhang2020universal}.
Floquet techniques have been experimentally demonstrated to simultaneously mitigate
decoherence due to longitudinal relaxation and pure dephasing
\cite{huang2020engineering, mundada2020floquet}.
Numerical QOC techniques have been adapted to mitigate longitudinal relaxation
by modeling master equations \cite{rembold2020introduction} and employing
Monte Carlo style quantum trajectories \cite{abdelhafez2019gradient}.
Additionally, efforts have been made to incorporate experimental feedback
into optimization \cite{huang2020engineering}. Comprehensive tools that interleave characterization
and pulse design are under active development \cite{wittler2020integrated}.

>>>>

\section{Introduction}
Quantum optimal control (QOC) techniques are a class of optimization
algorithms for accurately and efficiently manipulating quantum systems.
Early techniques were proposed for nuclear magnetic resonance experiments
\cite{khaneja2005optimal}, and applications now include superconducting
circuits \cite{heeres2017implementing,
  huang2020engineering, leng2019robust, leung2017speedup, xu2020nonadiabatic},
neutral and ionized atoms \cite{van2016optimal}, nitrogen-vacancy centers in
diamond \cite{rembold2020introduction}, and Bose-Einstien condensates
\cite{sorensen2018quantum}. For quantum computation
QOC techniques are employed to achieve high fidelity gates
while adhering to experimental constraints.
The decision variables of the optimization problem are the time-dependent control
parameters, particular to the quantum system, that govern its evolution.
Experimental errors may cause the system evolution to deviate from that predicted in
optimization, hampering performance.
The field of classical control theory has developed robust control techniques
to encode experimental errors in optimization objectives, improving
experimental performance \cite{Zhou97,Morimoto00,Manchester18}.
In this work we employ robust control techniques to mitigate
realistic system parameter deviations and decoherence that arise when controlling
a superconducting fluxonium qubit.

Analytic and numerical techniques for QOC have been remarkably successful in
designing high fidelity gates \todo{citations needed}. 
While these methods are able to predict system behavior with high accuracy, they tend 
to be sensitive to experimental errors such as parameter drift, noise, and finite control 
resolution. 
% In addition to the high
% simulated fidelities which can be obtained with these techniques, it is desirable for the
% resultant quantum state trajectory to be insensitive to experimental errors
% such as parameter drift, noise, and finite control resolution. Quantum state
% trajectories that are insensitive to these effects are said to be robust.
% In line with this objective, 
In order to address these issues,
recent work has sought to mitigate decoherence
and errors due to parameter deviations
Considering the dynamical and geometric phases of quantum state
trajectories has led to analytic methods for mitigating
errors due to parameter deviations and pure dephasing
\cite{han2020experimental, merrill2014progress, xu2020nonadiabatic, zhang2020universal}.
Floquet techniques have been experimentally demonstrated to simultaneously mitigate
decoherence due to longitudinal relaxation and pure dephasing
\cite{huang2020engineering, mundada2020floquet}.
Numerical QOC techniques have been adapted to mitigate longitudinal relaxation
by modeling master equations \cite{rembold2020introduction} and employing
Monte Carlo style quantum trajectories \cite{abdelhafez2019gradient}.
Additionally, efforts have been made to incorporate experimental feedback
into optimization \cite{huang2020engineering}. Comprehensive tools that interleave characterization
and pulse design are under active development \cite{wittler2020integrated}.

### 2020-11-20 paper/2background.tex
minor edits

Quantum optimal control seeks the control
parameters that minimize a functional $J(u(t))$.
In the simplest case the functional is
the infidelity between the inital state evolved
to the final knot point and the target state
$J = 1 - {\lvert \braket{\psi_{f} \lvert \psi_{N}(u(t))} \rvert}^{2}$
In general $J$ is a linear combinaion of cost functions on the state, e.g.
forbidden-state occupation, as well as
cost functions on the controls, e.g. the norm of the control amplitudes
\cite{leung2017speedup}. Typical quantum optimal control
algorithms employ automatic differentiation
to compute first order information for the functional $\nabla_{u} J(u)$.
This information is used in a first-order optimizer to minimize $J$ with respect to $u$.
This scheme lends itself to projected gradient methods
which restrict the optimization to the constraint manifold
\cite{clarkson2010coresets, hauswirth2016projected}.

Alternatively, the QOC problem can be formulated as a trajectory optimization problem 
and solved using any of the state-of-the-art solvers developed by the robotics community.
Trajectory optimization problems are typically of the following form: 
\begin{mini}[2]
    {x_{0:N},u_{0:N-1}}{\ell_f(x_N) + \sum_{k=0}^{N-1} \ell(x_k,u_k) }{}{}
    \addConstraint{x_{k+1} = f(x_k,u_k)}
    \addConstraint{g_k(x_k,u_k)}{\leq 0}
    \addConstraint{h_k(x_k,u_k)}{=0}
    \label{opt:discrete_trajopt},
\end{mini}
where $\ell_f$ and $\ell$ are the final and stage cost functions, $x_k \in \R^n$ and
$u_k \in \R^m$ are the state and input control variables, $f(x_k,u_k)$ is the discrete
dynamics function, and $g_k(x_k,u_k)$ and $h_k(x_k,u_k)$ are the inequality
and equality constraints, potentially including initial and final conditions,
at time step $k$.

@@

ALTRO uses iterative LQR (iLQR) \cite{Li2004a} as the internal solver of an augmented 
Lagrangian method (ALM). iLQR solves an unconstrained trajectory optimization problem 
using a backward Riccati recursion to derive a closed-loop linear feedback law about the current 
trajectory. By simulating the system forward with the feedback law, the trajectory is 
brought closer to the (locally) optimal trajectory. DDP-based solvers such as iLQR
are popular since they are very computationally efficient, are always dynamically feasible,
and provide a closed-loop control policy about the optimal trajectory for free; however,
standard implementations have no ability to deal with nonlinear equality and inequality 
constraints. ALM handles constraints by successively solving unconstrained minimization 
problems of the form:
\begin{mini}[2]
    {z}{f(z) + \lambda^T c(z) + \half c(z)^T I_\mu c(z)}{}{}
    \label{opt:alm}
\end{mini}
where $f(z)$ is the objective function, $c(z) : \R^n \mapsto \R^p$ is the constraint 
function, $\lambda \in \R^p$ are the Lagrange multipliers, and $I_\mu$ is a diagonal matrix
of penalty weights, whose magnitude depend on whether the constraint is active or inactive.
For ALTRO, $f(z)$ is the objective function of \eqref{opt:discrete_trajopt}, $c(z)$ is the 
concatenation of $g_k$ and $h_k$, and $z$ is the concatenation of the states and controls 
across all time steps. After minimizing \eqref{opt:alm} using iLQR, the penalty terms and 
Lagrange multipliers are updated, and the process repeats until convergence.

>>>>

Quantum optimal control seeks the control
parameters that minimize a functional $J(u(t))$.
In the simplest case the functional is
the infidelity between the inital state evolved
to the final knot point and the target state
$J = 1 - {\lvert \braket{\psi_{f} \lvert \psi_{N}(u(t))} \rvert}^{2}$
In general $J$ is a linear combinaion of cost functions on the state, e.g.
forbidden-state occupation, as well as
cost functions on the controls, e.g. the norm of the control amplitudes
\cite{leung2017speedup}. Typical quantum optimal control
algorithms employ automatic differentiation
to compute gradients of the functional $\nabla_{u} J(u)$,
which can easily be used to implement first-order optimization methods
\cite{clarkson2010coresets, hauswirth2016projected}.

Alternatively, the QOC problem can be formulated as a trajectory optimization problem 
and solved using a variety of specialized solvers developed by the robotics community \cite{Schulman13,Tedrake16,Hereid2017FROST,howell2019altro}.
Trajectory optimization problems are typically of the following form: 
\begin{mini}[2]
    {x_{0:N},u_{0:N-1}}{\ell_f(x_N) + \sum_{k=0}^{N-1} \ell(x_k,u_k) }{}{}
    \addConstraint{x_{k+1} = f(x_k,u_k)}
    \addConstraint{g_k(x_k,u_k)}{\leq 0}
    \addConstraint{h_k(x_k,u_k)}{=0}
    \label{opt:discrete_trajopt},
\end{mini}
where $\ell_f$ and $\ell$ are the final and stage cost functions, $x_k \in \R^n$ and
$u_k \in \R^m$ are the state and input control variables, $f(x_k,u_k)$ is the discrete
dynamics function, and $g_k(x_k,u_k)$ and $h_k(x_k,u_k)$ are the inequality
and equality constraints, potentially including initial and final conditions,
at time step $k$.

@@

ALTRO uses iterative LQR (iLQR) \cite{Li2004a} as the internal solver of an augmented 
Lagrangian method (ALM). iLQR solves an unconstrained trajectory optimization problem 
using a backward Riccati recursion to derive a closed-loop linear feedback law about the current 
trajectory. By simulating the system forward with the feedback law, the trajectory is 
brought closer to the (locally) optimal trajectory. DDP-based solvers such as iLQR
are popular since they are very computationally efficient, are always dynamically feasible,
and provide a locally stabilizing closed-loop control policy about the optimal trajectory. However,
standard implementations have no ability to deal with nonlinear equality and inequality 
constraints. ALM handles constraints by successively solving unconstrained minimization 
problems of the form:
\begin{mini}[2]
    {z}{f(z) + \lambda^T c(z) + \half c(z)^T I_\mu c(z)}{}{}
    \label{opt:alm}
\end{mini}
where $f(z)$ is the objective function, $c(z) : \R^n \mapsto \R^p$ is the constraint 
function, $\lambda \in \R^p$ is a Lagrange multiplier estimate, and $I_\mu$ is a diagonal matrix
of penalty weights, $\mu$, whose magnitudes depend on whether the corresponding constraint is active or inactive.
For ALTRO, $f(z)$ is the objective function of \eqref{opt:discrete_trajopt}, $c(z)$ is the 
concatenation of $g_k$ and $h_k$, and $z$ is the concatenation of the states and controls 
across all time steps. After minimizing \eqref{opt:alm} using iLQR, the Lagrange multiplier estimates are updated according to,
\begin{equation}
	\lambda \gets \lambda + \mu c(z) ,
\end{equation}
the penalty terms are updated, and the process repeats until convergence.

### 2020-11-20 paper/3fluxonium.tex
minor edits

In addition to imposing optimization constraints that
reflect physical limitations of the apparatus, we impose
constraints that improve the experimental realization of the control pulse.
To ensures gates may be concatenated arbitrarily without
inducing AWG ringing due to high-frequency transitions,
we require $a(t = 0) = a(t = t_{N}) = 0$.
Furthermore, we require $\int_{0}^{t_{N}} a(t) dt = 0$. This
constraint ensures the pulse has zero net flux, mitigating
the hysteresis ubiquitous in flux bias lines,
\cite{battistel2019fast, krantz2019quantum, zhang2020universal}.
We require $-0.5 \textrm{GHz} \le a(t) \le 0.5 \textrm{GHz}$
to ensure the two-level approximation \eqref{eq:hamiltonian}
remains valid. Additionally, we require that each gate achieves
the desired state transition $\ket{\psi_{N}} = \ket{\psi_{f}}$.
In addition to these constraints we penalize the norm
of the first and second derivatives of the flux amplitude to
ensure its smoothness, and in doing so mitigate AWG ringing.
The optimization is performed over the second derivative of the flux amplitude
$\pdv*[2]{a}{t}$ which is contained in the
augmented control vector. The first derivative
$\pdv*{a}{t}$, proportional $a$, and integral $\int a$
flux amplitude terms
are contained in the augmented state vector. They are obtained from
the second derivative of the flux amplitude by
integration in the dynamics function.
Both the zero net flux and target quantum state constraint
are then handled by ensuring the target augmented state is
reached $x_{N} = x_{f}$, the constraint functional 
\todo{not sure what the convention is in your community, but for us a functional is a 
function of functions (or trajectories). This would just be a function. Also, shouldn't 
this just be $x_N = x_f$ and $\abs{x_k} \le b$? Why are you squaring the functions?}
 takes the form
$c_{k} = (x_{k} - x_{f})^{T}(x_{k} - x_{f})$.
The equality and inequality constraints on $a$ are handled
with a bound constraint which takes the form
$c_{k} = (x_{k} - b)^{T}(x_{k} - b)$ if $\lvert x_{k} \rvert \ge b$
and $0$ otherwise.
In addition the norm of the quantum state
is constrained to ensure discretization error is not
exploited $c_{k} = {\lvert \braket{\psi_{k} \lvert \psi_{k}} \rvert}^{2} - 1$.
\todo{is this last constraint supposed to be an inequality constraint?}

>>>>

In addition to imposing optimization constraints that
reflect physical limitations of the apparatus, we impose
constraints that improve the experimental realization of the control pulse.
To ensure gates may be concatenated arbitrarily without
inducing AWG ringing due to high-frequency transitions,
we require $a(t = 0) = a(t = t_{N}) = 0$.
Furthermore, we require $\int_{0}^{t_{N}} a(t) dt = 0$. This
constraint ensures the pulse has zero net flux, mitigating
the hysteresis ubiquitous in flux bias lines,
\cite{battistel2019fast, krantz2019quantum, zhang2020universal}.
We require $-0.5 \textrm{GHz} \le a(t) \le 0.5 \textrm{GHz}$
to ensure the two-level approximation \eqref{eq:hamiltonian}
remains valid. Additionally, we require that each gate achieves
the desired state transition $\ket{\psi_{N}} = \ket{\psi_{f}}$.
In addition to these constraints we penalize the norm
of the first and second derivatives of the flux amplitude to
ensure its smoothness, and in doing so mitigate AWG ringing.
The optimization is performed over the second derivative of the flux amplitude
$\pdv*[2]{a}{t}$ which is contained in the
augmented control vector. The first derivative
$\pdv*{a}{t}$, proportional $a$, and integral $\int a$
flux amplitude terms
are contained in the augmented state vector. They are obtained from
the second derivative of the flux amplitude by
integration in the dynamics function.
Both the zero net flux and target quantum state constraint
are then handled by ensuring the target augmented state is
reached $x_{N} = x_{f}$, the constraint functional 
\todo{not sure what the convention is in your community, but for us a functional is a 
function of functions (or trajectories). This would just be a function. Also, shouldn't 
this just be $x_N = x_f$ and $\abs{x_k} \le b$? Why are you squaring the functions?}
 takes the form
$c_{k} = (x_{k} - x_{f})^{T}(x_{k} - x_{f})$.
The equality and inequality constraints on $a$ are handled
with a bound constraint which takes the form
$c_{k} = (x_{k} - b)^{T}(x_{k} - b)$ if $\lvert x_{k} \rvert \ge b$
and $0$ otherwise.
In addition the norm of the quantum state
is constrained to ensure discretization error is not
exploited $c_{k} = {\lvert \braket{\psi_{k} \lvert \psi_{k}} \rvert}^{2} - 1$.
\todo{is this last constraint supposed to be an inequality constraint?}
